### 2023-02-08 paper.tex
SNR and constellation wip

\documentclass[12pt]{report} % add 'draft' to suppress figures
% Double-spaced, single-column format for IEEE journal submission
% \documentclass[12pt,journal,draftclsnofoot,onecolumn]{IEEEtran} 
\usepackage{wrapfig,booktabs,fancyhdr,amsmath,amsfonts,tabularx,numprint}
\usepackage{cite,bm,bbm,amssymb,amsthm,url,multirow,times,enumitem,comment}
\usepackage{mathtools,siunitx,balance,tikz,adjustbox,graphicx,array}
\usepackage[font=footnotesize]{subcaption}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage[colorlinks=true, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }
 \usepackage[acronym,nomain,nonumberlist]{glossaries}
 \makeglossaries

 \newcommand{\vb}{\boldsymbol}
\newcommand{\vbh}[1]{\hat{\boldsymbol{#1}}}
\newcommand{\vbc}[1]{\check{\boldsymbol{#1}}}
\newcommand{\vbb}[1]{\bar{\boldsymbol{#1}}}
\newcommand{\vbt}[1]{\tilde{\boldsymbol{#1}}}
\newcommand{\vbs}[1]{{\boldsymbol{#1}}^*}
\newcommand{\vbd}[1]{\dot{{\boldsymbol{#1}}}}
\newcommand{\abs}[1]{\left|{#1}\right|}
\newcommand{\by}{\times}
\newcommand{\tr}{\mathsf{T}}
\newcommand{\sfrac}[2]{\textstyle \frac{#1}{#2}}
\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}
\newcommand{\sinc}{\text{sinc}}
\newcommand{\define}{\triangleq}
%\newcommand{\define}{\coloneqq}
\newcommand{\cnr}{C/N_0}
\newcommand{\sgn}{\text{sgn}}
\renewcommand{\Re}{\mathbb{R}}
\renewcommand{\Im}{\mathbb{I}}
\newcommand{\E}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\brm}[1]{\bm{\text{#1}}}
\newcommand{\round}[1]{\ensuremath{\lfloor#1\rceil}}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}

% Macros for this paper
\newcommand{\Fc}[1]{\ensuremath{F_{\text{c}#1}}}
\newcommand{\Fs}{\ensuremath{F_\text{s}}}
\newcommand{\Fsr}{\ensuremath{F_\text{sr}}}

\newacronym{PL}{PL}{Path loss}
\newacronym{LOS}{LOS}{Line of sight}
\newacronym{NLOS}{NLOS}{Non-line of sight}
\newacronym{BW}{BW}{Bandwidth}
\newacronym{iid}{i.i.d.}{independent and identically-distributed}
\newacronym{CLT}{CLT}{Central Limit Theorem}
\newacronym{GRV}{GRV}{Gaussian random variable}
\newacronym{SNR}{SNR}{signal-to-noise ratio}
\newacronym{PSD}{PSD}{power spectral density}
\glsaddall

% \topmargin = 0 mm 
% \oddsidemargin = -1 mm 
% \evensidemargin = -1 mm
% \headheight = 0 mm 
% \headsep = 8 mm 
% \textheight = 220 mm 
% \textwidth =170 mm 
% \parindent = 0 mm
% \parskip = 4 mm

\begin{document}
%%% ----------------------------------------------------------Front Matter
\title{An Exegesis of Ontological Hermeneutics}

\tableofcontents
\chapter{Path Loss Models}

\gls{PL} models are formulas which allow for modeling the received
power $P_r$ given the transmit power $P_t$, distance $d$, and signal carrier
wavelength $\lambda$. The resulting \gls{PL} is due to signal propagation
through the environment.

\section{Friss Formula}
For a transmitter and receiver spaced $d$ distance apart, the received power
$Pr$ follows:

\begin{align}
  \label{eq:general_PrPt}
  P_r = P_t \frac{1}{4\pi d} A_e
\end{align}

where $A_e = \frac{\lambda}{4 \pi}$ for an omnidirectional antenna and a signal
with a wavelength $\lambda$. Substituting $A_e$ in (\ref{eq:general_PrPt})
provides the Friis Formula, which applies for a wireless channel in a vacuum.
Further adding terms $G_t$ and $G_r$ for the transmit and receive antenna gains,
we reach the generalized Friis formula.

\begin{align}
  \label{eq:friss_formula}
  P_r = \frac{G_t G_r \lambda ^2 P_t}{(4\pi d)^2}
\end{align}

\section{Path Loss Exponent Model}
Since wireless channels exist in environments without a vacuum, a generalized
model is needed to cover a variety of cases. A generalized such model is an
exponential model using the received power formula below.

\begin{align}
  \label{eq:general_exp_PL}
  P_r = P_t G_t G_r K \left[\frac{d_0}{d}\right]^\alpha
\end{align}

The parameter $d_0$ is a unit normalization parameter, usually set to $1$ m. The
parameter $\alpha$ is the \gls{PL} exponent. The parameter $K$ is the nominal
\gls{PL} at $1$ m. Notice when $K = [\frac{\lambda}{4 \pi}]^2$ defined as $K_0$
and $\alpha = 2$, the formula reveals the Friis Formula from
(\ref{eq:friss_formula}).

\chapter{Multipath Effect Modeling}
Other than the natural landscape of the environment, there are other effects
that may further increase \gls{PL}.

\section{Blocking}
Blocking allows for a multi-slope \gls{PL} model. Essentially, the \gls{PL}
model has multiple definitions, and follows each one with a certain probability.
This usually translates to following a certain \gls{PL} for a \gls{LOS} path,
and following another model for the \gls{NLOS} path. A common model for this is
the Exponential Blocking model

\begin{align}
  \label{eq:exp_blocking}
  P_r = \left\{ \begin{array}{ll}
                  PL_{LOS}  & \text{w.p.} ~~~~~~ e^{-d/\beta}  \\
                  PL_{NLOS} & \text{w.p.} ~~~ 1 - e^{-d/\beta} \\
                \end{array}
  \right.
\end{align}

where $\beta$ acts as the mean distance before blocking occurs (i.e. by a building), and the $PL_{LOS}$ and $PL_{NLOS}$ terns are functions for \gls{PL} models.

\subsection*{Blocking example}
Imagine a scenario in which a \gls{LOS} path is in free space and the \gls{NLOS}
is modeled as an exponential \gls{PL} with an exponent $\alpha = 2.5$ and a
distance of $d = 100$ between the transmitter and receiver. The blocking
exponent $\beta = 25$. What does the $SNR = P_r/P_t$ model finalize to if $G =
  G_t G_r = 1$, $K_{\text{LOS}} = K_{\text{NLOS}} = K_0 = -40$ dB and $d_0 = d_1 =
  1$.
\\
\\
Answer:

\begin{align}
  SNR = \frac{P_r}{P_t} = \left\{ \begin{array}{ll}
                                    K_0 \left[\frac{1}{100}\right]^2     & \text{w.p.} ~~~~~~ e^{-100/25}  \\
                                    K_0 \left[\frac{1}{100}\right]^{2.5} & \text{w.p.} ~~~ 1 - e^{-100/25} \\
                                  \end{array}
  \right. \nonumber
  = \left\{ \begin{array}{ll}
              -80 \text{ dB} & \text{w.p.} ~ 0.0183 \\
              -90 \text{ dB} & \text{w.p.} ~ 0.9817 \\
            \end{array}
  \right. \nonumber
\end{align}

\section{Shadowing}
Shadowing is attenuation caused by objects between the transmitter and receiver
that reduce the received signal's power. It is modeled as random and does not
depend on the distance of the wireless communication link. By modeling shadowing
we account for the constructive and destructive interference cause by the
transmitted signal reflecting, diffracting and scattering, the components of
which are called multipath components.

The shadowing effect is modeled as lognormal. The received power equation from a
\gls{PL} model is simply multiplied with a new variable $\mathcal{X}$, where
$x_{\text{dB}} = 10 log_{10}{\mathcal{X}} \sim
  \mathcal{N}(0,\sigma_{\text{dB}}^2)$.

Alternatively, some use a model where $P_r = P_t \Psi$ where $ \psi_{\text{dB}}
  = 10 log_{10}{\Psi}  \sim \mathcal{N}(PL_{\text{dB}}(d),\sigma_{\text{dB}}^2)$
for some \gls{PL} model.

\subsection*{Shadowing example}
Imagine a scenario in which we wish to evaluate Wi-Fi coverage for a desired
range of $100$ m, with a minimum $SNR = 5$ dB and outage constraint at $1\%$ for
the $5$ GHz frequencies. The \gls{PL} is modeled as exponential with $\alpha =
  3$, $K = K_0 = -50$ dB (for $F_c \approx 5$ GHz), no TX and RX gain, a noise
power of $P_n = -100$ dBm and a shadowing standard deviation of
$\sigma_{\text{dB}} = 6$. What is the mimum transmition power to meet the outage
constraint?
\\
\\
Answer:

\begin{align}
  SNR             & = \frac{P_r}{P_t} \geq 5 \text{ dB} \nonumber                                       \\
  P_r             & \geq -95 \text{ dB} \nonumber                                                       \\
  P_r             & = P_{t,\text{dB}} + K_{0,\text{dB}} + x_{\text{dB}} - \alpha 10 log_{10}d \nonumber \\
  P_{t,\text{dB}} & \geq -95 + 50 - x_{\text{dB}} + 60 \nonumber                                        \\
  P_{t,\text{dB}} & \geq 15 - x_{\text{dB}} \nonumber
\end{align}

Since $x_{\text{dB}}$ is normally distributed, we simply use the CDF of a
Gaussian distribution to evaluate transmit power that would allow for the
required outage constraint of $0.01$:

\begin{align}
  P(P_{t,\text{dB}} \geq  15 - x_{\text{dB}}) = 1 - 0.01 \nonumber \\
  15 - P_{t,\text{dB}} = \sigma_{\text{dB}}Q^{-1}(0.99) \nonumber  \\
  P_{t,\text{dB}} \geq 29 \text{ dBm} \nonumber
\end{align}

\chapter{Multipath Channel Models}
\label{multipath_ch_model}
A transmitted signal $s(t)$ propagates through an environment in multiple paths.
Neglecting noise, the received signal is simply a sum of attenuated, delayed,
and phase shifted versions of the original transmitted signal. The different
paths a signal takes to get to a receiver results in a sum with patterns of high
or low destructive interference, an effect called multipath fading.
\section{Multipath Model}
We model the channel $c(\tau,t)$ as a time-varying ($t$) impulse response function evaluated at $t - \tau$. As with any system, the received signal is modeled as a convolution of $c(\tau,t)$ with the transmitted signal $s(t)$.

A way to interpret the parameters is as follows:
\begin{itemize}
  \item $t$ is the time an impulse response is observed
  \item $\tau$ is the offset from the observation time in which the impulse was transmitted
\end{itemize}

The traditional impulse response of a time-invariant system can be expressed as
thinking of the channel's impulse response as periodic $c(\tau,t + T)$ where $T
  = -t$ such that  $c(\tau,0) = c(\tau)$, which is the impulse response at time
$0$.

Whether \gls{NLOS} received signals are received in a large spread of time
compared to the \gls{LOS} component or not, depends on the inverse of the
transmitted signal's \gls{BW}. This spread is called the delay spread $T_m$.
Depending on whether it is large or small changes the model significantly, so it
is explored as Narrowband and Wideband fading.

\section{Narrowband Fading Model}
Under this regime, the delay spread $T_m$ is significantly smaller than the inverse of the baseband \gls{BW} of the signal $B_{bb}$, $T_m << B^{-1}_{bb}$. Essentially, this model assumes that all paths have no delay as mentioned in Section \ref{multipath_ch_model}, but still have random phases. For a received signal $r(t)$, the following model is assumed,

\begin{align}
  r(t) = \operatorname{Re} \Biggl\{ s(t) e^{j2\pi f_c t} \left(\sum_i \alpha_i(t) e^{-j \phi_i(t)}\right) \Biggr\}
\end{align}
where $f_c$ is the carrier frequency, $\alpha_i$ are the attenuation on each path $i$ with their respective phase offset $\phi_i$.

For large $N(t)$ (time dependent number of multipath components) if $\alpha_i$ and $\phi_i$ are \gls{iid}, then by the \gls{CLT}, the real and imaginary components of the baseband equivalent received signal are approximately joint \glspl{GRV}. If some path is larger in amplitude, like the \gls{LOS} path, then the paths are no longer \gls{iid} so the \gls{CLT} no longer holds and the baseband components are not joint \gls{GRV}.

TODO Autocorrelation
\chapter{Channel Reliability}

\section{Signal Modulation}

TODO

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/constellations.pdf}
  \caption{Constellation schemes where Quadrature and In-Phase possible discrete options are displayed in terms of their energy.}
  \label{fig:constellations}
\end{figure*}

\begin{table}[htbp]
  \centering
  \caption{}
  \begin{tabular}{lcccc}                                                                                                                   \\
    \cmidrule(r){1-5}  & MPAM                           & MPSK                       & MQAM                         & FSK     \\
    \midrule $E_s$     & $\frac{d_{min}^2}{12}(M^2-1)$  & $d^2$                      & $\frac{d_{min}^2}{6}(M^2-1)$ & $d^2/2$ \\
    \midrule $d_{min}$ & $\sqrt{\frac{12E_s}{M^2 - 1}}$
                       & $\sqrt{E_s}$                   & $\sqrt{\frac{6}{M^2 - 1}}$ & $\sqrt{2 E_s}$                         \\

    \bottomrule
  \end{tabular}
  \label{tab:win-space-results}
\end{table}

\section{Signal-to-noise ratio}

If the complex noise $n(t) \sim \mathcal{C}\mathcal{N}(0,2\sigma^2)$, meaning
each of the two channels is distributed as a zero-mean \gls{GRV} with variance
$\sigma$, this relates to the noise \gls{PSD} $N_0$ (in watts per hertz $W/Hz$)
as $\sigma^2 = N_0/2$. If the energy per dimension is $\bar{E_s} = E_s/2$, then
the \gls{SNR} is defined as:

\begin{align}
  \gamma_s = SNR = \frac{\bar{E_s}}{\sigma} = \frac{E_s}{2\sigma} = \frac{E_s}{N_0}
\end{align}

\Gls{SNR} can also be defined in terms of the power of the received signal $P_s$
over the power of the noise $P_n = N_0 B$, where $B = 1/T$ is the \gls{BW} of
the receiver. Note, the received signal power $P_s$ relates to the energy per
symbol as $P_s = E_s/T$ for a symbol time T.

\begin{align}
  \gamma_s = SNR = \frac{P_s}{P_n}= \frac{E_s/T}{N_0 B} = \frac{E_s}{N_0}
\end{align}

The \gls{SNR} also has an equivalent metric, the per-bit \gls{SNR} $Ebb-N_0$
expressed in terms of the energy per bit $E_b$ over the noise \gls{PSD} $N_0$.
The energy per bit is $E_b = E_s/\log _{2} M$. Also, define the bit-rate as $R =
  \frac{\log _{2} M}{T}$.

\begin{align}
  \gamma_b = Ebb-N_0 = \frac{E_b}{N_0} = \frac{ E_s/\log _{2} M}{N_0} = \frac{\gamma_s}{\log _{2} M} = \gamma_s \frac{B}{R}
\end{align}

\section{Probability of Error}
TODO

The probability of error $P_s$ is defined in a per-symbol basis. Each constellation has a different $P_s$, seen in the table below based on the Nearest Neighbor bound.

\begin{table}[htbp]
  \centering
  \caption{}
  \begin{tabular}{lcccc}                                                                                \\
    \cmidrule(r){1-3} & $P_s(\gamma_s)$                                  & $P_b(\gamma_b)$ \\
    \midrule MPAM     & $2(1-1/M)Q(\sqrt{\frac{3\gamma_s}{M^2-1}})$      &                 \\
    \midrule MPSQ     & $2Q(\sqrt{2\gamma_s}\sin(\pi/M))$                                  \\
    \midrule MQAM     & $4(1-1/\sqrt{M})Q(\sqrt{\frac{3\gamma_s}{M-1}})$                   \\
    \midrule FSK      & $(M-1)Q(\sqrt(\gamma_s/2))$                                        \\
    \\

    \bottomrule
  \end{tabular}
  \label{tab:constellation_pe}
\end{table}

Useful approximations for $0 \leq \gamma_s \leq 30$ dB:
\begin{itemize}
  \item MQAM $M \geq 4$ has $P_b(\gamma_s) \approxeq 0.2 e^{\frac{-1.5\gamma_s}{M-1}}$
  \item BPSK has $P_b(\gamma_s) = P_s(\gamma_s) \approxeq 2 e^{-1.5\gamma_s}$
\end{itemize}

\printglossary[type=\acronymtype,title=Acronyms]

% \bibliographystyle{ieeetran}
% \bibliography{ref}
\end{document}